---
output:
  html_document: default
  pdf_document: default
---
Presidential Campaign Analysis in Texas by Rohit MIttapalli
========================================================  

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(knitr)
library(dplyr)
library(gridExtra)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load the Data
elec_db <- read.csv('New Dataset.csv',header = TRUE)
```

The loaded dataset represents financial contributions to presidential campaigns
in Texas. The dataset contains the amount, date, election term of each
contribution, the name of the nominee contributed to, and the employer, name,
location, occupation of the contributor.

Using external data from the census.gov, population data of cities were added.
The gender and political leaning for each candidate was also added. This was
manually collected and joined using python scripts.

# Univariate Plots Section

To begin, We will study the typical donated amount.  

```{r echo=FALSE, warning=FALSE,Univariate_Plots , message=FALSE}
p1 <- ggplot(aes(x = contb_receipt_amt),data = elec_db) +
  geom_histogram(binwidth = 5) +
  scale_x_continuous(limits = 
                       quantile(elec_db$contb_receipt_amt,probs=c(.05, .95))) +
  ggtitle("Distribution of Contribution",subtitle = "0.05-.95% : Width = 5")

p2 <- ggplot(aes(x = contb_receipt_amt),data = elec_db) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = 
                       quantile(elec_db$contb_receipt_amt,probs=c(.05, .95))) +
  ggtitle("Distribution of Contribution",subtitle = "0.05-.95% : Width = 25")

p3 <- ggplot(aes(x = contb_receipt_amt),data = elec_db) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(limits = c(0,100)) +
  ggtitle("Distribution of Contribution",subtitle = "0-100$ Width = 1")

grid.arrange(p1,p2,p3)

```

We will also include a summary of across contribution amount

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(elec_db$contb_receipt_amt)
```


We see primarily that donations are made at incremental values: 25, 50, 100, 200,
250, and 300. The first two plots use the middle 95% of the data to remove
outlier and the last plot studies points of over plotting in the first two.

We will now look at the amount of contributions over time.


```{r echo=FALSE, message=FALSE, warning=FALSE}

elec_db$contrb_dates <- as.Date(elec_db$contb_receipt_dt,
                                "%d-%b-%y",origin="01-Jan-1960")
beg <- as.Date("2016-01-01")
end <- as.Date("2017-01-01")

p1 <- ggplot(aes(x=contrb_dates),data = elec_db) +
  geom_histogram(binwidth = 5) +
  scale_x_date(limits = c(beg,end),date_breaks = "2 month",
               date_labels = "%b-%y") +
  ggtitle("Distribution of Contribution Dates",subtitle = "Width = 5 days")

p2 <- ggplot(aes(x=contrb_dates),data = elec_db) +
  geom_histogram(binwidth = 15) +
  scale_x_date(limits = c(beg,end),date_breaks = "2 month", 
               date_labels = "%b-%y") +
  ggtitle("Distribution of Contribution Dates",subtitle = "Width = 15 days")

p3 <- ggplot(aes(x=contrb_dates),data = elec_db) +
  geom_histogram(binwidth = 30) +
  scale_x_date(limits = c(beg,end),date_breaks = "2 month",
               date_labels = "%b-%y") +
  ggtitle("Distribution of Contribution Dates",subtitle = "width = 30 days")

grid.arrange(p1,p2,p3)

```


We can see the number of contributions increasing over time and fluctuating
periodically over time. There seems to be peaks around March, July, and
November.

We will now look at a histogram of contribution city populations. To properly
see city sizes the x-axis has been scaled by the sqrt function.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x=Population),data = subset(elec_db,!is.na(Population))) +
  geom_histogram(binwidth = 50) +
  scale_x_sqrt() +
  ggtitle("Distribution of Contribution Population",
          subtitle = "width = 30 people")
  

```

We will also include a summary of across city population.

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(elec_db$Population)
```

There seems to be two types of cities those under 500 thousand residents and
those above. We will categorize cities greater than 500,000 as large cities and
those under as small.

We will now look into the proportion of big and small cities. We have
logarithmically scaled the y-axis.

```{r echo=FALSE, message=FALSE, warning=FALSE}
df <- data.frame(Size=c("Small City", "Big City"),
                Count=c(1026,11))

ggplot(data=df, aes(x=Size, y=Count)) +
  geom_bar(stat="identity") +
  scale_y_log10()
```

We see there are only 11 big cities and about a 1000 small ones. We will now
study the total number of contributions by large and small cities.

```{r echo=FALSE, message=FALSE, warning=FALSE}
df <- data.frame(Size=c("Small City", "Big City"),
                Count=c(sum(elec_db$Population < 500000,na.rm = TRUE), 
                        sum(elec_db$Population > 500000,na.rm = TRUE)))

df2 <- data.frame(Size=c("Small City", "Big City"),
                Count=c(sum(elec_db$Population < 500000,na.rm = TRUE)/1026, 
                        sum(elec_db$Population > 500000,na.rm = TRUE)/11))

p1 <- ggplot(data=df, aes(x=Size, y=Count)) +
  geom_bar(stat="identity") +
  ggtitle("Number of Contributions for Big and Small City",subtitle = "Raw")
p2 <- ggplot(data=df2, aes(x=Size, y=Count)) +
  geom_bar(stat="identity") +
    ggtitle("Number of Contributions for Big and Small City", 
            subtitle = "Normalized") 
    
grid.arrange(p1,p2)

```

Surprisingly although there are 100 times less big cities than small cities the
number of contributions are in the same magnitude.

We will now look into the candidates. We will start by examining the gender of
the candidates. We will take a look at the number of female and male candidates.

```{r echo=FALSE, message=FALSE, warning=FALSE}
df <- data.frame(dose=c("Male", "Female"),
                len=c(23, 3))
ggplot(data=df, aes(x=dose, y=len)) +
  geom_bar(stat="identity")
```

There are significantly fewer female candidates than male candidates.

We will now look at the political leaning of each candidate.

```{r echo=FALSE, message=FALSE, warning=FALSE}
df <- data.frame(Party=c("Republican","Democrat","Libertarian","Green"),
                Count=c(19,5,1,1))
ggplot(data=df, aes(x=Party, y=Count)) +
  geom_bar(stat="identity")
```

We see that there are significantly more Republican candidates than Democrat
candidates and only a single Green and Libertarian candidate.

We will now see the number of contributions for each gender.

```{r echo=FALSE, message=FALSE, warning=FALSE}
df <- data.frame(Gender=c("Female","Male"),
                Count=c(sum(elec_db$Gender == "F"),sum(elec_db$Gender == "M")))

df2 <- data.frame(Gender=c("Female","Male"),
                Count=c(sum(elec_db$Gender == "F")/3,
                        sum(elec_db$Gender == "M")/23))

p1 <- ggplot(data=df, aes(x=Gender, y=Count)) +
  geom_bar(stat="identity") +
  ggtitle("Number of Contributions by Gender",subtitle = "Raw")


p2 <- ggplot(data=df2, aes(x=Gender, y=Count)) +
  geom_bar(stat="identity") +
    ggtitle("Number of Contributions by Genders",
            subtitle = "Normalized by number of candidates")


grid.arrange(p1,p2)
```

We see that despite there are significantly fewer female candidates, female
candidates receive about two-thirds the number of contributions as male
candidates.

We will now look at the number of contributions for each party.

```{r echo=FALSE, message=FALSE, warning=FALSE}
df <- data.frame(Party=c("Republican","Democrat","Libertarian","Green"),
                Count=c(sum(elec_db$Party == "Republican"),
                        sum(elec_db$Party == "Democrat"),
                        sum(elec_db$Party == "Libertarian"),
                        sum(elec_db$Party == "Green")))

df2 <- data.frame(Party=c("Republican","Democrat","Libertarian","Green"),
                Count=c(sum(elec_db$Party == "Republican")/19,
                        sum(elec_db$Party == "Democrat")/5,
                        sum(elec_db$Party == "Libertarian"),
                        sum(elec_db$Party == "Green")))

p1 <- ggplot(data=df, aes(x=Party, y=Count)) +
  geom_bar(stat="identity") +
  ggtitle("Number of Contributions by Party",subtitle = "Raw")


p2 <- ggplot(data=df2, aes(x=Party, y=Count)) +
  geom_bar(stat="identity") +
    ggtitle("Number of Contributions by Party",
            subtitle = "Normalized by number of candidates")


grid.arrange(p1,p2)
```

We see that despite there being significantly fewer number of democratic
candidates, democratic candidates received more contributions than all other
parties. Green and Libertarian parties do not have a significant number of
contributions. 

# Univariate Analysis

### What is the structure of your dataset?

The dataset is structured as a series of events. Each event is a contribution
to presidential candidates described by many variables. The dataset primarily
ranges across 2016.

### What is/are the main feature(s) of interest in your dataset?

The main features in the dataset include gender, political leniency,
contribution rate over time and city population. 

We see that the number of female and democratics candidates are fewer than the 
number of male and republican candidates. Despite there being significantly
fewer candidates, female candidates make two-thirds of the number of contributions
of male candidates. Similarly, there are fewer democratic candidates than
republican candidates however democratic candidates received more contributions. 

This idea is also expanded to big and small cities. Large cities (categorized by
a population greater than 500,000 people) are 100 times fewer than small cities
but have two thirds of contributions of small cities.

In general, a minority still has a significant number of contributions.

Other features is that the number of contributions seem to follow a cycle with
peaks in March, July and November and lows in June and September.

The number of donations are often made at incremental values with the majority
sub 100.

### What other features in the dataset do you think will help support your \

investigation into your feature(s) of interest?

I think if more data across different years were included, more political trends
could be derived. It would also be interesting to see if the income of donators
correlated with contribution.

### Did you create any new variables from existing variables in the dataset?

I created new variables in the sense that I normalized certain data. For example
for the number of contribution of candidates per gender, I normalized them by
the number of 3 candidates.

I also added 3 columns of data including city population, candidate gender and
political affiliation.

I also converted the date column to be of datetime objects. 

### Of the features you investigated, were there any unusual distributions? \

The number of contributions over time has an unusual distribution with three
large peaks. The population of cities also had an unusual distribution due to
large cities like Houston and Dallas. To accommodate that, I set it to a square
root scale. For the number of cities above and below 500,000, I scaled the y
axis by log 10.

# Bivariate Plots Section

We will look at average, median, and total contribution by time. For this plot
I had to extract the month and year from the data.

```{r echo=FALSE, warning=FALSE, message=FALSE}

elec_db$Month <- as.Date(cut(elec_db$contrb_dates,
  breaks = "month"))

month_elec_db_groups <- group_by(elec_db, Month)
elec_db_groups <- summarise(month_elec_db_groups,
                          elec_count_mean = mean(contb_receipt_amt),
                          elec_count_median = median(contb_receipt_amt),
                          elec_count_sum = sum(contb_receipt_amt),
                          n = n())
elec_db_groups <- arrange(elec_db_groups, Month)

p1 <- ggplot(aes(x=Month , y = elec_count_mean), data = elec_db_groups) +
  geom_line() +
  ggtitle("Average Contribution over Time") +
  scale_x_date(limits = c(beg,end),date_breaks = "2 month",
               date_labels = "%b-%y")


p2 <- ggplot(aes(x=Month , y = elec_count_median), data = elec_db_groups) +
  geom_line() +
  ggtitle("Median Contribution over Time") +
  scale_x_date(limits = c(beg,end),date_breaks = "2 month", 
               date_labels = "%b-%y")


p3 <- ggplot(aes(x=Month , y = elec_count_sum), data = elec_db_groups) +
  geom_line() +
  ggtitle("Total Contribution over Time") +
  scale_x_date(limits = c(beg,end),date_breaks = "2 month", 
               date_labels = "%b-%y")


p4 <- ggplot(aes(x=Month , y = n), data = elec_db_groups) +
  geom_line() +
  ggtitle("Number of Contributions over Time") +
  scale_x_date(limits = c(beg,end),date_breaks = "2 month", 
               date_labels = "%b-%y")



grid.arrange(p1,p2,p3,p4)
```

We see that the average and median contribution over time seems to be relatively
steady. However there seems to be low total contribution around April and July 
and a peak near February and October. We can contribute that to a lower number of
contributions which also peaks in February and October and lowers in April and 
July.

We will know look at average, median, and total contribution by term.

```{r echo=FALSE, warning=FALSE, message=FALSE}

term_elec_db_groups <- group_by(elec_db, election_tp)
elec_db_term <- summarise(term_elec_db_groups,
                          elec_count_mean = mean(contb_receipt_amt),
                          elec_count_median = median(contb_receipt_amt),
                          elec_count_sum = sum(contb_receipt_amt),
                          n = n())

p1 <- ggplot(aes(x=election_tp , y = elec_count_mean), data = elec_db_term) +
  geom_bar(stat = "identity") +
  ggtitle("Average Contribution over Time")

p2 <- ggplot(aes(x=election_tp , y = elec_count_median), data = elec_db_term) +
  geom_bar(stat = "identity") +
  ggtitle("Median Contribution over Time") 

p3 <- ggplot(aes(x=election_tp , y = elec_count_sum), data = elec_db_term) +
  geom_bar(stat = "identity") +
  ggtitle("Total Contribution over Time") 

p4 <- ggplot(aes(x=election_tp , y = n), data = elec_db_term) +
  geom_bar(stat = "identity") +
  ggtitle("Number of Contributions over Time") 



grid.arrange(p1,p2,p3,p4)

```

We first note that there is not significant data for most terms and the only
terms we can study are O2016 and P2016. We see that O2016 has a significantly
higher average and median contribution but less number of contributions which
results in a total less contribution over time.

We will now look at average contribution and contribution based on gender.

```{r echo=FALSE, warning=FALSE, message=FALSE}
p1 <- ggplot(aes(x = contb_receipt_amt),data = elec_db) +
  geom_histogram(binwidth = 10) +
  facet_wrap(~elec_db$Gender)+
  scale_x_continuous(limits = 
                       quantile(elec_db$contb_receipt_amt,probs=c(.05, .95))) +
  ggtitle("Distribution of Contribution by Gender",
          subtitle = "0.05-.95% : Width = 10") 

p2 <- ggplot(aes(x=Gender, y=contb_receipt_amt),data = elec_db) +
  geom_boxplot() +
    coord_cartesian(ylim = 
                      quantile(elec_db$contb_receipt_amt,probs=c(.05, .95))) +
    ggtitle("Distribution of Contribution by Gender",subtitle = "0.05-.95%")


grid.arrange(p1,p2)

```

We see that female candidates typically receive more smaller donations than male
candidates but receive less donations greater than 100$. Overall the average
donation for female candidates is less than the average contribution from males
although they are very close in range.

We will now look at average contribution and contribution based on political
party.

```{r echo=FALSE, warning=FALSE, message=FALSE}
p1 <- ggplot(aes(x = contb_receipt_amt),data = elec_db) +
  geom_histogram(binwidth = 10) +
  facet_wrap(~elec_db$Party)+
  scale_x_continuous(limits = 
                       quantile(elec_db$contb_receipt_amt,probs=c(.05, .95))) +
  ggtitle("Distribution of Contribution by Party",
          subtitle = "0.05-.95% : Width = 10") +
  scale_y_sqrt()

p2 <- ggplot(aes(x=Party, y=contb_receipt_amt),data = elec_db) +
  geom_boxplot() +
    coord_cartesian(ylim = 
                      quantile(elec_db$contb_receipt_amt,probs=c(.05, .95))) +
    ggtitle("Distribution of Contribution by Party",subtitle = "0.05-.95%")


grid.arrange(p1,p2)

```

We see at first that green and libertarian contributions are limited in number
and resultantly have a relatively small sample size to draw from. In general,
democrat nominees have more donations less than 100$ than their republican
counterparts whereas republican candidates have more donations above 100$.
Overall the average donation for Democrats is less than all other parties.

We will now look at city population by candidate's political party.

```{r echo=FALSE, warning=FALSE, message=FALSE}


ggplot(aes(x=Party, y=Population),data = elec_db) +
  geom_boxplot() +
    coord_cartesian(ylim = quantile(elec_db$Population,
                                    probs=c(.05, .95),na.rm = TRUE)) +
    ggtitle("Distribution of City Population by Party",subtitle = "0.05-.95%")



```

We see interestingly that average city population for democrats and green party
donations is greater than 500,000 whereas republican and libertarian is
significantly less. 

We will now look at city population by gender.

```{r echo=FALSE, warning=FALSE, message=FALSE}


ggplot(aes(x=Gender, y=Population),data = elec_db) +
  geom_boxplot() +
    coord_cartesian(ylim = 
                  quantile(elec_db$Population,probs=c(.05, .95),na.rm = TRUE)) +
    ggtitle("Distribution of City Population by Gender",subtitle = "0.05-.95%")



```

We see that the average city population voting for female candidates is
significantly higher than male candidates.

We will now look at average contribution by city population.

```{r echo=FALSE, warning=FALSE}

ggplot(aes(x = Population, y = contb_receipt_amt),data = elec_db) +
  geom_point(alpha = 0.01) +
    scale_y_continuous(limits = 
                        quantile(elec_db$contb_receipt_amt,probs=c(.05, .95))) +
    ggtitle("Contribution Amount by City Population",subtitle = "0.05-.95%") +
  geom_smooth()


```

We see that for the most part the average contribution is relatively  constant
across city population. There does not seem to be a correlation between
population and contribution amount.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \

We see that the average and median contribution by time is realtively constant
however, there is a periodicity in total contributions which can be attributed
to a fluctuating number of contributions.

We cannot draw many conclusions about contribution by terms since there is not
significant data about each term.

We also see that female candidates receive a smaller average contribution than
their male counterparts. We also note that contributions towards female
candidates are typically from larger cities than their male candidates.

This is also similar for political affiliation. There is not significant data
to draw conclusions for the libertarian and green parties. The democratic party
receives smaller average contributions than the republican party and typically
draw from larger cities than the republican party.

It is important to note that the parallel between gender and political party is
likely influenced from the fact that leading democratic candidate was Hillary 
Clinton: a female whereas the Republican candidates were primarily male. 

We also saw that contribution by city population was relatively constant.

### What was the strongest relationship you found?

The three strongest relationships were between city population and political
lenience, city population and gender, number of contributions and time.

# Multivariate Plots Section

We first look at contribution amount by city population with party colors.


```{r echo=FALSE, warning=FALSE}
ggplot(aes(x = contrb_dates, y = Population),data = elec_db) +
  scale_y_sqrt() + 
  scale_x_date(limits = c(beg,end),date_breaks = "2 month",
               date_labels = "%b-%y") +
  geom_point(aes(color = Party),alpha = 0.01) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  scale_color_manual(values=c("#0826ad", "#2a7025", "#e2dd6f","#ad0707")) +
    ggtitle("Average City Population over Time by Party") 
```

This plot showed a strong relationship between number of contributions and
political affiliation. We see that there are bands of colors across all cities.
We also note that larger cities while influenced by the bands are slightly
more consistent.

In the next plot, we look at contribution amount by time and political party.

```{r echo=FALSE, warning=FALSE, message=FALSE}

p1 <- ggplot(aes(x = contrb_dates, y = contb_receipt_amt),data = elec_db) +
  scale_y_log10() + 
  scale_x_date(limits = c(beg,end),date_breaks = "2 month",
               date_labels = "%b-%y") +
  geom_point(aes(color = Party),alpha = 0.01) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  scale_color_manual(values=c("#0826ad", "#2a7025", "#e2dd6f","#ad0707")) +
    ggtitle("Average Contribution over Time by Party")


p2 <- ggplot(aes(x=contrb_dates , y = contb_receipt_amt), data = elec_db) +
  geom_point(alpha = 0.01) +
  ggtitle("Average Contribution over Time by Party") +
  scale_x_date(limits = c(beg,end),date_breaks = "2 month",
               date_labels = "%b-%y") +
  facet_wrap(~Party) +
  scale_y_log10()+ guides(colour = guide_legend(override.aes = list(alpha = 1)))

grid.arrange(p1,p2)

```

We see the bands being reciprocated across this plot as well. The main band of
interesting is between May and June where Republican donation seemingly stopped.
Another interesting feature is that after Aug 2016 republican donation for
values under 100$ essentially stopped. For the most part democratic donation has
been constant but after November they also stopped.

We will now look at the same thing with gender.

```{r echo=FALSE}
ggplot(aes(x = contrb_dates, y = Population),data = elec_db) +
  scale_y_sqrt() + 
  scale_x_date(limits = c(beg,end),date_breaks = "2 month",
               date_labels = "%b-%y") +
  geom_point(aes(color = Gender),alpha = 0.03) +
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
    scale_color_manual(values=c("#e845e5", "#1e8dc9"))

```

This plot seems to align with party lines (Female to Democrat and Male to
Republican). However the bands of higher population seem to fall as susceptible
to the bands unlike party plots.

We will now look at contribution amount by time and gender.

```{r echo=FALSE, warning=FALSE, message=FALSE}

p1 <- ggplot(aes(x = contrb_dates, y = contb_receipt_amt),data = elec_db) +
  scale_y_log10() + 
  scale_x_date(limits = c(beg,end),date_breaks = "2 month",
               date_labels = "%b-%y") +
  geom_point(aes(color = Gender),alpha = 0.01) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
    scale_color_manual(values=c("#e29ce1", "#55e0e0"))

p2 <- ggplot(aes(x=contrb_dates , y = contb_receipt_amt), data = elec_db) +
  geom_point(alpha = 0.01) +
  ggtitle("Average Contribution over Time") +
  scale_x_date(limits = c(beg,end),date_breaks = "2 month", 
               date_labels = "%b-%y") +
  facet_wrap(~Gender) +
  scale_y_log10() + guides(colour =guide_legend(override.aes = list(alpha = 1)))

grid.arrange(p1,p2)

```

Again, we see similar patterns to party lines. However, the gap in donations from
May to June for Republican candidates is only prominent for donations above 100$.
Female candidates received the same amount of donations relatively consistent 
with a major drop after November.

Similar to republican candidates, male candidates after November faced a serious
drop in donations under a 100$.

To explore these bands further we will look at number of contribution by time
and by gender.

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(aes(x = contrb_dates),data = elec_db) +
  geom_histogram() +
  facet_wrap(~Party) +
      scale_x_date(limits = c(beg,end),date_breaks = "2 month",
                   date_labels = "%b-%y") +
  ggtitle("Number of Contributions by Time and Party")

```

Here we see the source of these bands and the fluctuation on donations over time
in the data. We see that there is a much more significant dip in the donations
over time in June and September for the Republican party than in the Democratic
side. However there is a peak in March and October, however the October peak
for Democrats, is significantly more substantial than the Republicans.

We will now look at contribution over time by city population.

```{r echo=FALSE, warning=FALSE, message=FALSE}

ggplot(aes(x = contrb_dates, y = contb_receipt_amt),data = elec_db) +
  scale_y_sqrt() +
    scale_x_date(limits = c(beg,end),date_breaks = "2 month",
                 date_labels = "%b-%y") +
  geom_point(aes(color = elec_db$Population)) +
  scale_color_gradientn(colours = rainbow(5))+
  ggtitle("Contributions Amount by Time and City Population")





```

The homogeneous mixture of colors shows us that despite population and time,
contribution amount stays relatively constant.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \

The main relationships we observed were: 

    Contribution Amount by City Population by Party
    Contribution Amount by City Population by Gender
    Contribution Amount by Time by Party
    Contribution Amount by Time by Gender
    Contribution Amount by Time by City Population

### Were there any interesting or surprising interactions between features?

The most interesting elements of these plots were clear bands of Republican and
Democratic donations. We see that these parallel the number of donations over
time by party. We see clear drops for both in June and September. However,
the republican party faces a significantly bigger drop in contribution in these
lows. There are peaks in donations in March and October; however the Democratic
peak in October is much more substantial.

Other interesting components included that after the month of August Republican
donation under 100$ virtually stopped. 

We also saw that there was not a clear correlation between contributed amount
and city population or city population donation and time.

------

# Final Plots and Summary

### Plot One

```{r echo=FALSE, warning=FALSE, message=FALSE}

p1 <- ggplot(aes(x = contrb_dates, y = contb_receipt_amt),data = elec_db) +
  ylab("Contribution Amount (USD)") +
  xlab("Contribution Dates") +
  scale_y_log10() + 
  scale_x_date(limits = c(beg,end),date_breaks = "2 month",
               date_labels = "%b-%y") +
  geom_point(aes(color = Party),alpha = 0.01) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  scale_color_manual(values=c("#0826ad", "#2a7025", "#e2dd6f","#ad0707")) +
    ggtitle("Average Contribution over Time by Party")


p2 <- ggplot(aes(x=contrb_dates , y = contb_receipt_amt), data = elec_db) +
  geom_point(alpha = 0.01) +
  ylab("Contribution Amount (USD)") +
  xlab("Contribution Dates") +
  ggtitle("Average Contribution over Time by Party") +
  scale_x_date(limits = c(beg,end),date_breaks = "2 month",
               date_labels = "%b-%y") +
  facet_wrap(~Party) +
  scale_y_log10()+ guides(colour = guide_legend(override.aes = list(alpha = 1)))


grid.arrange(p1,p2)

```

### Description One

This plot is a plot of average contribution over time and by party. The most
interesting part of this plot are bands of Democratic and Republican dominance.
Especially between May and June 2016, there are very few Republican donations.
Also after August 2016 there are essentially no republican donations under a 
100$.

This plot made me consider the number of contributions over time for different
parties. This plot is shown below.

### Plot Two

```{r echo=FALSE, warning = FALSE}
ggplot(aes(x = contrb_dates),data = elec_db) +
  geom_histogram() +
  facet_wrap(~Party) +
      scale_x_date(limits = c(beg,end),date_breaks = "2 month",
                   date_labels = "%b-%y") +
  ggtitle("Number of Contributions by Time and Party")+
    ylab("Number Of Contributions") + xlab("Contribution Date")
```

### Description Two

There are a few interesting components to this plot. First of for both
republican and democratic parties there are maximums in March, July and October
and lows in June and September. This was also reflected in the total
contribution over time. However interestingly, republican donations are far
more affected by these lows. In June and September there are very few Republican
donations at all. Also in November 

### Plot Three

```{r echo=FALSE, warning = FALSE}


ggplot(aes(x=Party, y=Population),data = elec_db) +
  geom_boxplot() +
    coord_cartesian(ylim = 
                  quantile(elec_db$Population,probs=c(.05, .95),na.rm = TRUE)) +
    ggtitle("Distribution of City Population by Party",subtitle = "0.05-.95%")
 

```

### Description Three

This last plot shows the distribution of city population by party. Most
donations to democrats and green parties come from cities of big populations
whereas republican donations often come from smaller cities. 

------

# Reflection

The project studied political contributions in Texas. The insights showed that
in many cases minorities (Female Candidates, Democratic Nominees, and Number of
Large Cities) had a significant number and total contribution as their majority
counterparts. Furthermore, the analysis gave insight to political lows and highs
and major contributing elements of donations. Surprisingly, it showed an
interesting correlation between city population and political lenience.

Over time it was interesting how certain explorations led to further questions.
For example seeing the bands in city population donations over time by party
led me to question total number of contribution over time by party. This 
exploration led to more questions and a better understanding.

Before I began studying this data, I started with all the elements provided by
the Federal Election Commission. I quickly learned that this data simply wasn't
sufficient. I required more information about the candidate's affiliated party
to say more about the nature of the donation. Moreover, I required data about
city size to tell more about nature of the donator. Gather this external data
and joining it with that provided resulted in a more holistic understanding of
the data.

In the future, I hope to gather more data across years to see if these trend 
hold. Often throughout this exploration, political party and gender aligned.
This is most likely due to the leading democratic candidate being Female. With
data from more years, I suspect this correlation will not be as cut and dry.
In comparison, libertarian and green party do not have much data. Collecting
across a few years may provide the data to properly analyze the parties.